---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# emWeightedCPI Package

<!-- badges: start -->

<!-- badges: end -->

The goal of emWeightedCPI package is to automate a proposed multivariate weighted indexing scheme for price data. The multivariate weighted consumer price index (MVWCPI) generates weights for both base and current years, respectively, from base and current year price datasets using communality obtained after performing factor analysis. The package takes base and current year price data as inputs. This approach is convenient since it does not require incurring additional cost for conducting household expenditure survey to generate CPI for the determination of inflation figures. However, the package can still generate the Laspeyres’ CPI and inflation using base year expenditure data only. More information
[here](http://dx.doi.org/10.9734/bpi/mono/978-81-19315-32-1).

## Installation

You can install the development version of mvWeightedCPI from [GitHub](https://github.com/) like so:

``` r
install.packages("devtools")
devtools::install_github("JC-Ayimah/emWeightedCPI")
```

## Package Functions

### mvw_cpi function

The function that calculates the multivariate weighted indices is the
`mvw_cpi` function. It requires only one argument *`(data)`*; the price
data from which to calculate the indices.

```{r mvw_cpi function}
library(emWeightedCPI)
#create an arbitrary price data
mypriceData <- data.frame(x1=runif(50, 9.9, 13.7),
                    x2=rnorm(50, 10.9, 2.1),
                    x3=runif(50, 12.2, 15),
                    x4=runif(50,19.4, 24), # base year prices
                    y1=runif(50, 26, 30),
                    y2=runif(50, 31, 38.9),
                    y3=runif(50, 28.2, 33.1),
                    y4=runif(50, 51.8, 60) # current year prices
)
#call the function and specify the price data
mvw_cpi(data = mypriceData)
```


### mvw_inflation function

The `mvw_inflation` function uses the determined multivariate weighted indices from the `mvw_cpi` to 
further calculate inflation. Four index values are generated from the `mvw_cpi` function and you will 
have to specify which index to use in the `mvw_inflaion` function. The function takes two arguments 
*`index`* and *`data`* . the *`index`* argument takes one of four possible values(index); *`'fisher'`*, 
*`'paashe'`*,*`'laspeyres'`* and *`'drobish'`*. The *`data`* argument requires the price dataset from
which the inflation is to be determined.

```{r mvw_inflation function}
#calculating inflation using Fishers index
mvw_inflation(index = 'fisher', data = mypriceData)
```

Fishers' is the ideal index for calculating inflation. The rest of the
indices generates a warning after calculating inflation

```{r ideal index}
#generates warning after calculating inflation
mvw_inflation(index = 'drobish', data = mypriceData)
```

### eb_cpi Function
This function also calculates consumer price index. Whiles `mvw_cpi` uses multivariate weights,
the `eb_cpi` determines CPI from expenditure based expenses. The function takes in two inputs, a price data and an expenditure data.The index calculated from this function is the *`'Laspeyres index'`*
```{r eb_cpi function}
#pick base year prices from mypriceData
n_vec <- apply(mypriceData[, 1:4], 2, mean)

#combine the output above into a dataframe with two columns and same 
#number of rows as number of price items to create expenditure data
myexpData <- cbind.data.frame(item = names(n_vec), price = (unname(n_vec)))

#calculate CPI 
eb_cpi(price_data = mypriceData, expenditure_data = myexpData)
```
### eb_inflation Function
Like the `mvw_inflation` function, the `eb_inflation` function also calculates inflation based on 
the index calculated from the `eb_cpi` function. The function takes the same arguments specified
in the `eb_cpi` function.
```{r eb_inflation function}
eb_inflation(mypriceData, myexpData)
```

## Package Datasets
The package contains some price datasets
`(empirical_Data1, empirical_Data2 and expenditure_Data)` [more information on them in the
help documentation], that can be used with the above functions. The data is
Lazy loaded into your session once you import the package with
`library()` so you can just view the data by calling the name of the
data

```{r package dataset1}
# inspect empirical_Data1 provided in package
head(empirical_Data1)
dim(empirical_Data1)
# calculate mvw_cpi indices and mvw_inflation using empirical_Data1
mvw_cpi(data = empirical_Data1)
mvw_inflation(index='fisher', data = empirical_Data1)
```
```{r package dataset2}
# inspect empirical_Data2 provided in package
head(empirical_Data2)
dim(empirical_Data2)
# calculate indices and inflation using empirical_Data2
mvw_cpi(data = empirical_Data2)
mvw_inflation(index='fisher', data = empirical_Data2)
```
```{r package dataset3}
#inspect expenditure_Data provided in The package
head(expenditure_Data)
dim(expenditure_Data)
#calculate eb_cpi and eb_inflation using empirical_Data1 and expenditure_Data
eb_cpi(empirical_Data1, expenditure_Data)
eb_inflation(empirical_Data1, expenditure_Data)
```
