---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mvWeightedCPI Package

<!-- badges: start -->

<!-- badges: end -->

The goal of mvWeightedCPI is to generate multivariate weighted indices
for a price data. mvWeightedCPI generates these weights from base and
current year price datasets. You provide the needed data and
mvWeightedCPI works behind the scene to generate these weights. This
approach is convenient since it does not require incurring additional
cost for conducting household expenditure survey to generate CPI for the
determination of inflation figures.

## Installation

You can install the development version of mvWeightedCPI from [GitHub](https://github.com/) like so:

``` r
install.packages("devtools")
devtools::install_github("JC-Ayimah/mvWeightedCPI")
```

## Package Functions

### mvw_cpi function

The function that calculates the multivariate weighted indices is the
`mvw_cpi` function. It requires only one argument *`(data)`*; the price
data from which to calculate the indices.

```{r mvw_cpi function}
library(mvWeightedCPI)
#create an arbitrary price data
myData <- data.frame(x1=runif(50, 9.9, 13.7),
                    x2=rnorm(50, 10.9, 2.1),
                    x3=runif(50, 12.2, 15),
                    x4=runif(50,19.4, 24),
                    y1=runif(50, 26, 30),
                    y2=runif(50, 31, 38.9),
                    y3=runif(50, 28.2, 33.1),
                    y4=runif(50, 51.8, 60)
)
#call the function and specify the price data
mvw_cpi(data = myData)
```


### mvw_inflation function

The package also contains another function; `mvw_inflation` that uses
the determined multivariate weighted indices from the `mvw_cpi` to
further calculate inflation. Four index values are generated from the
`mvw_cpi` function and you will have to specify which index to use in
the `mvw_inflaion` function. The function takes two arguments *`index`*
and *`data`* . the *`index`* argument takes one of four possible
values(index); *`'fisher'`*, *`'paashe'`*,*`'laspeyres'`* and
*`'drobish'`*. The *`data`* argument requires the price dataset from
which the inflation is to be determined.

```{r mvw_inflation function}
#calculating inflation using Fishers index
mvw_inflation(index = 'fisher', data = myData)
```

Fishers' is the ideal index for calculating inflation. The rest of the
indices generates a warning after calculating inflation

```{r ideal index}
mvw_inflation(index = 'drobish', data = myData)
```

## Package Datasets
The package contains some price datasets
`(empirical_Data1 and empirical_Data2)` [more information on them in the
help documentation], that can be used with the function. The data is
Lazy loaded into your session once you import the package with
`library()` so you can just view the data by calling the name of the
data

```{r package dataset1}
# inspect empirical_Data1 provided in package
head(empirical_Data1)
dim(empirical_Data1)
# calculate indices and inflation using empirical_Data1
mvw_cpi(data = empirical_Data1)
mvw_inflation(index='fisher', data = empirical_Data1)
```
```{r package dataset2}
# inspect empirical_Data2 provided in package
head(empirical_Data2)
dim(empirical_Data2)
# calculate indices and inflation using empirical_Data2
mvw_cpi(data = empirical_Data2)
mvw_inflation(index='fisher', data = empirical_Data2)
```
