---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mvWeightedCPI

<!-- badges: start -->
<!-- badges: end -->

The goal of mvWeightedCPI is to generate multivariate weighted indices for a price data. 
mvWeightedCPI generates these weights from base and current year price datasets. You 
provide the needed data and mvWeightedCPI works behind the scene to generate these 
weights. This approach is convenient since it does not require incurring additional cost 
for conducting household expenditure survey to generate CPI for the determination of 
inflation figures.

## Installation

You can install the development version of mvWeightedCPI like so:

``` r
# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
```

## Example Usage

The package already contains price data that it uses to calculate the weighted indices.
The datasets contained in the package are two; `empirical_Data1` and `empirical_Data2`. 
You can just call the function without specifying the data and it will randomly pick one of the two datasets to determine the weights. 

```{r example}
library(mvWeightedCPI)
#calling the function without any arguments
mvw_cpi()
```

When you specify which dataset to use, the output indices always remain the same. The indices might change if you call the function more than once without specifying the data due to the random selection of the price data.
```{r}
# choose which dataset to use for calculating weights
mvw_cpi(data = 'empirical_Data2')
```
```{r}
#output might change when function is called more than once without specifying data 
#firs call
mvw_cpi()

#second call
mvw_cpi()

#third call
mvw_cpi()
```

The package also contains another function; `inflation` that uses the determined multivariate weighted indices from the `mvw_cpi` to further calculate inflation. 
Four index values are generated from the `mvw_cpi` function. The function has only one argument *`weights`* which takes four possible values(index); *`'fisher'`*, *`'paashe'`*,*`'laspeyres'`* and *`'drobish'`*. You can however specify only one possible index
```{r}
#calculating inflation using Fishers index
inflation(weights = 'fisher')
```

